### 2017-05-09 Paper Drafts/Draft1/Draft1.tex
First Draft: Functional Representation

While monitoring bat populations provides useful information, understanding the root causes and effects of what is observed requires that the natural history of extant bat species is also well understood. Traits, such as call structure or body size, exhibited by particular bat species are linked to the bats interactions with its environment. \cite{aldridge1987morphology} Existing fossil records are of limited use in inferring the traits exhibited by ancestral bats, particularly with respect to echolocation calls for example. The reconstruction of ancestral traits relies heavily on the comparative analysis\cite{felsenstein2004inferring} of extant bat species. Thus, statistical data science techniques may be particularly useful for inferring the evolutionary dynamics and reconstructing ancestral states of echolocation in bats.

Previous studies of bat echolocation calls for both classification \cite{walters2012continental} and ancestral reconstruction \cite{collen2012evolution} examined features of the call extracted from the spectrogram of the call. These, somewhat arbitrary, call features relied on significant domain knowledge to ensure they were sensibly selected and used. More recently however, general techniques for the classification of acoustic signals have been developed. \cite{stathopoulos2014bat} \cite{damoulas2010bayesian} These methods do not require, but can be augmented by, domain knowledge. General techniques for ancestral reconstruction of function-valued traits, such as speech sounds or echolocation calls, have been proposed \cite{group2012phylogenetic}. The study of bat echolocation calls offers an opportunity to examine the efficacy of these techniques.

@@

This general approach to evolutionary inference for function-valued traits is implemented here for a set of bat echolocation calls. Our goal in doing so is twofold. These techniques had previously been considered in the context of modelling the evolution of human speech sounds in language. \cite{group2012phylogenetic} It is hoped that by applying these methods in the simpler context of the evolution of bat echolocation calls that progress can be made towards resolving methodological problems with these methods. For example, what is the most appropriate method of representing the information contained in an acoustic signal?

@@

A functional data object is generated when repeated measurements of some process are taken along a continuous scale, such as time. \cite{ramsay2006functional} These measurements can be thought of as representing points on a curve varies gradually and continuously. In the context of phylogenetics, these functional data objects are called function-valued traits. \cite{meyer2005up}

@@

The discrete Fourier transform of the signal \(\{\tilde{x}_{l'm'}^{S'}(\cdot)\}\) is given by\[\tilde{X}_{lm}^S(k) = \sum_{n = 0}^{N_{lm}^S - 1} \tilde{x}_{lm}^{S}(n) e^{-i2\pi kn / N_{lm}^S }.\] The energy spectral density of this signal is the magnitude of the Fourier transform and so the log spectral density per second (in decibel) is therefore \[\tilde{\mathcal{E}}_{lm}^S(k) = 10 \log_{10} \left(\tilde{X}_{lm}^S(k)\right) / f_S.\]

Bat echolocation calls can be considered as function-valued traits. A trait which is repeatedly

>>>>

While monitoring bat populations provides useful information, understanding the root causes and effects of what is observed requires that the natural history of extant bat species is also well understood. Traits, such as call structure or body size, exhibited by particular bat species are linked to the bats interactions with its environment. \cite{aldridge1987morphology} Existing fossil records are of limited use in inferring the traits exhibited by ancestral bats, particularly with respect to echolocation calls. The reconstruction of ancestral traits relies heavily on the comparative analysis\cite{felsenstein2004inferring} of extant bat species. Thus, statistical data science techniques may be particularly useful for inferring the evolutionary dynamics and reconstructing ancestral states of echolocation in bats.

Previous studies of bat echolocation calls for both classification \cite{walters2012continental} and ancestral reconstruction \cite{collen2012evolution} examined features of the call extracted from the call spectrogram. These, somewhat arbitrary, call features relied on significant domain knowledge to ensure they were sensibly selected and used. More recently however, general techniques for the classification of acoustic signals have been developed. \cite{stathopoulos2014bat} \cite{damoulas2010bayesian} These methods do not require, but can be augmented by, domain knowledge. General techniques for ancestral reconstruction of function-valued traits, such as speech sounds or echolocation calls, have been proposed \cite{group2012phylogenetic}. The study of bat echolocation calls offers an opportunity to examine the efficacy of these techniques.

@@

This general approach to evolutionary inference for function-valued traits is implemented here for a set of bat echolocation calls. Our goal in doing so is twofold. These techniques had previously been considered in the context of modelling the evolution of human speech sounds in language. \cite{group2012phylogenetic} It is hoped that by applying these methods in the simpler context of the evolution of bat echolocation calls that progress can be made towards resolving methodological problems. For example, how do we extend these methods to more realistic models of evolution?

@@

A functional data object is generated when repeated measurements of some process are taken along a continuous scale, such as time. \cite{ramsay2006functional} 
These measurements can be thought of as representing points on a curve that varies gradually and continuously. In the context of phylogenetics, these functional data objects are called function-valued traits. \cite{meyer2005up}

@@

The discrete Fourier transform of the signal \(\{\tilde{x}_{l'm'}^{S'}(\cdot)\}\) is given by\[\tilde{X}_{lm}^S(k) = \sum_{n = 0}^{N_{lm}^S - 1} \tilde{x}_{lm}^{S}(n) e^{-i2\pi kn / N_{lm}^S }.\] 
The energy spectral density of this signal is the magnitude of the Fourier transform and so the log energy spectral density per second (in decibel) is estimated by \[\tilde{\mathcal{E}}_{lm}^S(k) = 10 \log_{10} \left( \frac{|\tilde{X}_{lm}^S(k)|f_s}{N_{lm}^S}\right).\]
The log energy spectral density estimate for the signal \(\tilde{x}_{lm}^S(\cdot)\), \(\tilde{\mathcal{E}}_{lm}^S(\cdot)\) is now considered to be a noisy estimate of the log energy spectral density for species \(S\), denoted \(\mathcal{E}^S(\cdot)\). The log energy spectral density is a periodic function of frequency which describes the energy of a signal at each frequency on the interval \([0, \frac{f_S}{2}]\). \(\tilde{\mathcal{E}}_{lm}^S(\cdot)\) has been mapped to the decibel scale and also scaled according to the length, in time, of \(\tilde{x}_{lm}^S(\cdot)\).\cite{antoniou2006digital} We have now mapped each echolocation call on the same scale and \(\tilde{\mathcal{E}}_{lm}^S(\cdot)\) is now comparable to \(\tilde{\mathcal{E}}_{l'm'}^{S'}(f)\) at frequency \(f\). In order for these representations to be considered as function-valued traits however, \(\tilde{\mathcal{E}}_{lm}^S(\cdot)\) must be smoothed such the call representation in the frequency domain varies gradually and continuously.

The smoothed log energy spectral density is estimated by smoothing splines where \[
\mathcal{E}_{lm}^S(f) = \arg \min_{\mathcal{E}_{lm}^S(\cdot)} \int_{0}^{\frac{f_S}{2}}  \{ \tilde{\mathcal{E}}_{lm}^S(\cdot) - \mathcal{E}_{lm}^S(f) \}^2 df + \lambda \int_{0}^{\frac{f_S}{2}} \{ {\mathcal{E}_{lm}^{S}}'' (f)\}^2 df.
\]
The smoothing parameter \(\lambda\) is chosen using a generalised cross-validation procedure.\cite{friedman2001elements} \cite{citeR}

We now have a functional representation of each bats echolocation call where the pairs of observations \(\{f, \mathcal{E}_{lm}^S(f)\}\) and \(\{f, \mathcal{E}_{l'm'}^(S')(f)\}\) are directly comparable. The function-valued traits can now be modelled for evolutionary inference.
